The Plasma Physics of Dilute Ultrarelativistic \epm\ Pair Beams I:\\ Linear Theory

### 2016-04-26 oblique.tex
fixed up appendix

We discuss the solution to Equation (\ref{eq:dispersion kinetic}) in Appendix \ref{sec:solution kinetic}.  The associate growth rate for the kinetic beam plasma instability is  
\begin{equation}
\Gamma \simeq - \Gamma_0
\frac{\pi \gamma_w^2 \gph^3 m (\vph-v_{bz})}{4 \gamma_b T K_2(m/T)}
\left[
\left( \cG'^2 + 2\cG' + 2 \right) 
+
\frac{\gamma_b^2 v_{bx}^2}{2\cG^2} \frac{m^2}{T^2} \left(2 \cG' + 2\right)
\right]
\frac{e^{-\cG'}}{\cG'^3}\,.
\label{eq:OGgen}
\end{equation}

In Figure \ref{fig:ObliqueGMZ}, we plot the maximum growth rate (maximized over $\vph$) as function of $v_{bx} = |k_{bx} v_b/k|$ and $T$.  From the figure, it is clear that the growth rate is maximized for $v_{bx}/v_b \gg 1/\gamma_b$.  Moreover the maximum growth rates vary little and are robust for broad range of $v_{bx}/v_b = k_{bx}/k$, i.e., large range in angles between beam and wavevector.  Hence, this suggests that we might estimate the maximal oblique growth rates by taking the limit as $v_{bx}\gg v_b/\gamma_b$ and maximizing over $\vph$.  This is further seen in Figure \ref{fig:OGgen}, we maximize over both $v_{bx}$ and $\vph$.  Here for relativistic beams, the maximum growth rate varies little with $T$, varying by less than $10\%$ between hot and cold beams.  Thus, we find: 
\begin{equation}
\Gamma_M
\simeq
\left\{
\begin{aligned}
& 0.38 \Gamma_0 & T\ll m\\
& 0.34 \Gamma_0 & T\gg m
\end{aligned}
\right.\,,
\end{equation}
for $\gamma_b\gtrsim10$, as seen in Figure \ref{fig:OGgen}.

@@

To compute the contribution from the beam term, we will reorient our coordinate system and define the $z'$-axis along the wavevector, $\kvec$.  In this case we have the beam taking on a non-$z'$ component,  $\vel_b = v_{bz'}\zphat+v_{bx}\xphat$.   
This frame moves with a velocity, $\vel_{\rm ph} = \omega_k/k \zphat$.  With an eye toward computing the residue in (??), we define $p_{z'}=\gz\vz\epsilon_\perp$, $\epsilon=\gz\epsilon_\perp$, and
$\epsilon_\perp=\sqrt{m^2+p_\perp^2}$ is the perpendicular energy.  In this case, we can rewrite the beam distribution function as
\begin{eqnarray}
F_B &=& \frac{m_e c^2}{4\pi\gamma_B k_B T K_2(m_ec^2/k_B T )m_e^3c^3} \exp\left(-\frac {\gamma_B(E - v_{B,z'} p_{z'} - v_{B,x'} p_{x'})} {k_B T}\right)\\
%F_b &=& \frac{1}{4\pi\gamma_B m^2 T K_2(m/T)} \exp\left(-\frac {\gammabeam(E - \betabeamz p_{z'} - \betabeamx p_{x'}}{T}\right),\label{eq:dist1}\\
&=& \frac{\mu m_e^{-3}c^{-3}}{4\pi\gamma_B K_2(\mu)}  \exp\left(-\frac {\gamma_B\gz(c^2 - v_{B,z'}v_{z'})\epsilon_\perp)} {k_BT c^2}\right)\exp\left(\frac{\gamma_Bv_{B,x'} p_{x'}}{k_BT}\right),\label{eq:dist2}
\end{eqnarray}
Plugging the equation into (??), we find that the this involves the integral of 
\begin{equation}
\int \frac{k\partial F_b/\partial p_{z'}}{\omega - k_{z'}v_{bz'}}d^3p = i 2\pi R,
\end{equation}
where $R$ is the residue for $p_{z'}$ such that $v_{z'} = v_{\rm ph}$.  
   Here we find for the two elements that contribute to the pole:
\begin{equation}
\left.
\frac{\partial}{\partial p_{z'}} \left(\omega-k v_{bz'}\right)
\right|_{\rm pole}
=
\left.
- k \left(
  \frac{c^2}{\epsilon} - \frac{p_{z'}^2 c^4}{\epsilon^3}
\right)
\right|_{\rm pole}
=
- \frac{k c^2}{\gph^3 \epsilon_\perp}\,,
\end{equation}
and 
\begin{equation}
\begin{aligned}
\left.
k\frac{\partial F_b}{\partial p_{z'}}
\right|_{\rm pole}
%&=
&=
\left.
-\frac{k\gamma_B}{k_BT} \left( \frac{p_{z'} c^2}{\epsilon} - v_{B,z'}\right)
F_b
\right|_{\rm pole}\\
&=
-
\frac{
  k
  (\vph-v_{B,z'}) \mu m_e^{-3}c^{-3}}{
  4\pi k_BT K_2(\mu)
} \exp\left(-\frac {\gamma_B\gph(c^2 - v_{B,z'}v_{\rm ph})\epsilon_\perp)} {k_BT c^2}\right)\\
&\exp\left(\frac{\gamma_Bv_{B,x'} p_{x'}}{k_BT}\right).
\end{aligned}
\end{equation}
%where we have used
%\begin{equation}
%\left.
%\gamma_b(\epsilon-v_{bz} p_z -v_{bx} p_x)/T
%\right|_{\rm pole}
%=
%\gamma_b \gph ( 1-v_{bz}\vph ) \epsilon_\perp/T - \gamma_b v_{bx} p_x/T\,.
%\label{eq:Eopole}
%\end{equation}

@@

Thus, we define $p_{x'}'=\gamma_w(p_{x'}-w \epsilon_\perp)$ and $p_{y'}'=p_{y'}$ and find:
\begin{eqnarray}
%\epsilon'_\perp &=& \gamma_w(\epsilon_\perp - w p_x),\\
\epsilon_\perp &=& \gamma_w(\epsilon'_\perp + w p'_x),\\
d p_x d p_y &=& \frac{\epsilon_\perp}{\epsilon'_\perp} d p'_x d p'_y.
\end{eqnarray}
Inserting this into equation (\ref{eq:I-integral}), we find
\begin{equation}
\begin{aligned}
\kI &= 
\int d^2\!p'_\perp \frac{\epsilon_\perp^2}{\epsilon'_\perp} \exp\left(-\frac{\cG'\epsilon'_\perp}{k_B T}\right)\\
%&=
%\gamma_w^2 \int d^2\!p'_\perp \left(
 % \epsilon'_\perp + 2 w 
%\qc{p'_x}
%+ w^2 \frac{p'^2_x}{\epsilon'_\perp}
%\right) e^{-\cG'\epsilon'_\perp/m}\\
&=
\pi \gamma_w^2 \int_0^\infty dp'^2_\perp \left(
  \epsilon'_\perp + \frac{w^2}{2} \frac{p'^2_\perp}{\epsilon'_\perp}
\right) \exp\left(-\frac{\cG'\epsilon'_\perp}{k_B T}\right)\,,
\end{aligned}
\end{equation}
where $\cG'\equiv\cG/\gamma_w$. Note in the second line that we have used isotropy in $\bp_{\perp}'$ to eliminate terms linear in $\bp_{\perp}'$. 
% and in the second line we used the fact
%that  $\epsilon'_\perp e^{-\cG'\epsilon'_\perp/m}$ is isotropic in
%$\bp'_\perp$.  
Using the following integrals: 
\begin{equation}
\int_0^\infty dx \sqrt{1+x} e^{-a\sqrt{1+x}} = \frac{2 e^{-a}}{a^3}\left(a^2 + 2a + 2\right)
\end{equation}
and 
\begin{equation}
\int_0^\infty dx \frac{x}{\sqrt{1+x}} e^{-a\sqrt{1+x}}
=
\frac{4(a+1)}{a^3} e^{-a}\,,
\end{equation}
we find
\begin{equation}
\kI
=
\frac{2 \pi \gamma_w^2 (k_B T)^3}{\cG'^3} \left[
\left(\qc{\cG'}^2+2\qc{\cG'}+2\right)
+
\frac{w^2}{2}
\left(2\qc{\cG'}+2\right)
\right]\,\exp\left(-\frac{\cG'}{k_B T}\right).
\end{equation}
Inserting this into (\ref{eq:Rogen}) yields
\begin{equation}
R
=
\frac{
  \gph^3 m^2 (\vph-v_{bz}) 
}{
  2 T^2 K_2(m/T)
}
\gamma_w^2
\left[
\left(\cG'^2 + 2\cG' + 2\right)
+
\frac{\gamma_b^2 v_{bx}^2}{2\cG^2} \frac{m^2}{T^2} \left(2 \cG' + 2\right)
\right] \frac{e^{-\cG'}}{\cG'^3}
\end{equation}
and therefore,
\begin{equation}
\Gamma \simeq - \Gamma_0
\frac{\pi \gamma_w^2 \gph^3 m (\vph-v_{bz})}{4 \gamma_b T K_2(m/T)}
\left[
\left( \cG'^2 + 2\cG' + 2 \right) 
+
\frac{\gamma_b^2 v_{bx}^2}{2\cG^2} \frac{m^2}{T^2} \left(2 \cG' + 2\right)
\right]
\frac{e^{-\cG'}}{\cG'^3}\,.
\label{eq:OGgen}
\end{equation}

%The above differs from the beam plasma instability only due to the
%minor difference in the definition of $\cG'$ and the second term.  When
%$v_{bx}=0$, it is precisely the beam plasma instability growth rate.

@@

However, obtaining asymptotic expansions for the growth rate in the high and
low temperature limits that can be easily solved for the maximum
growth rate is difficult.  Even with the tremendous numerical insight
provided by Figure \ref{fig:ObliqueGMZ}, it is unclear how to make
significant progress.  However, some general remarks can be made.  It
appears at the maximum growth rate, we have to leading order:
\begin{equation}
\begin{gathered}
\vph\simeq\sqrt{2\delta Z_x}\,,\quad
v_{bx}=(1-\delta Z_x) v_b\,,\quad
v_{bz}\simeq\sqrt{2\delta Z_x}\,v_b\,,\quad\\
v_{bz}-\vph\simeq\left\{
\begin{aligned}
&\frac{\sqrt{T/m}}{\gamma_b} & T\ll m\\
&\frac{1}{2\gamma_b} & T\gg m
\end{aligned}
\right.\,,\quad
\gph\simeq1+\delta Z_x\,,\quad
\cG' \sim \frac{m}{T}\,,\\
\text{and}\quad w \sim 1\,,
\end{gathered}
\end{equation}
\qc{[can you add more details which equations you approximate under which conditions? e.g., I get $1/2\gamma_b^2$ for the hot beam case \ldots]}\\
where $\delta Z_x\equiv 1 - v_{bx}/v_b$.  Inserting these into Equation
(\ref{eq:OGgen}) suffices to show that there exists growth rates that
are $\propto \Gamma_0$, with 
\begin{equation}
\Gamma_M
\simeq
\left\{
\begin{aligned}
& 0.38 \Gamma_0 & T\ll m\\
& 0.34 \Gamma_0 & T\gg m
\end{aligned}
\right.\,,
\end{equation}
for $\gamma_b\gtrsim10$, as seen in Figure (\ref{fig:ObliqueGMZ}).

>>>>

We discuss the solution to Equation (\ref{eq:dispersion kinetic}) in Appendix \ref{sec:solution kinetic}.  The associate growth rate for the kinetic beam plasma instability is  
\begin{equation}
\Gamma \simeq - \Gamma_0
\frac{\pi \gamma_w^2 \gph^3 m (\vph-v_{B,z'})}{4 \gamma_b \mu K_2(\mu) \cG'^3}
\left[
\left( \cG'^2\mu^2 + 2\cG'\mu + 2 \right) 
+
\frac{\gamma_B^2 v_{B,x'}^2}{2\cG^2c^2} \left(2 \cG'\mu+ 2\right)
\right]
\exp(-\cG'\mu),
\end{equation}
where $\Gamma_0 \equiv \omega_p \gamma_B (n_b/n_t) (m_e v_B^2/k_B T)$ is the typical maximum growth rate.

In Figure \ref{fig:ObliqueGMZ}, we plot the maximum growth rate (maximized over $\vph$) as function of $v_{bx} = |k_{bx} v_b/k|$ and $T$.  From the figure, it is clear that the growth rate is maximized for $v_{bx}/v_b \gg 1/\gamma_b$.  Moreover the maximum growth rates vary little and are robust for broad range of $v_{bx}/v_b = k_{bx}/k$, i.e., large range in angles between beam and wavevector.  Hence, this suggests that we might estimate the maximal oblique growth rates by taking the limit as $v_{bx}\gg v_b/\gamma_b$ and maximizing over $\vph$.  This is further seen in Figure \ref{fig:OGgen}, we maximize over both $v_{bx}$ and $\vph$.  Here for relativistic beams, the maximum growth rate varies little with $T$, varying by less than $10\%$ between hot and cold beams.  Thus, we find: 
\begin{equation}
\Gamma_M
\simeq
\left\{
\begin{aligned}
& 0.38 \Gamma_0 & k_BT/m_ec^2\ll 1\\
& 0.34 \Gamma_0 & k_BT/m_ec^2\gg 1
\end{aligned}
\right.\,,
\end{equation}
for $\gamma_b\gtrsim10$, as seen in Figure \ref{fig:OGgen}.

@@

To compute the contribution from the beam term, we will reorient our coordinate system and define the $z'$-axis along the wavevector, $\kvec$.  In this case we have the beam taking on a non-$z'$ component,  $\vel_b = v_{bz'}\zphat+v_{bx}\xphat$.   
This frame moves with a velocity, $\vel_{\rm ph} = \omega_k/k \zphat$.  With an eye toward computing the residue in (??), we define $p_{z'}=\gz\vz\epsilon_\perp$, $\epsilon=\gz\epsilon_\perp$, and
$\epsilon_\perp=\sqrt{m^2+p_\perp^2}$ is the perpendicular energy.  In this case, we can rewrite the beam distribution function as
\begin{eqnarray}
F_B &=& \frac{m_e c^2}{4\pi\gamma_B k_B T K_2(m_ec^2/k_B T )m_e^3c^3} \exp\left(-\frac {\gamma_B(E - v_{B,z'} p_{z'} - v_{B,x'} p_{x'})} {k_B T}\right)\\
%F_b &=& \frac{1}{4\pi\gamma_B m^2 T K_2(m/T)} \exp\left(-\frac {\gammabeam(E - \betabeamz p_{z'} - \betabeamx p_{x'}}{T}\right),\label{eq:dist1}\\
&=& \frac{\mu m_e^{-3}c^{-3}}{4\pi\gamma_B K_2(\mu)}  \exp\left(-\frac {\gamma_B\gz(c^2 - v_{B,z'}v_{z'})\epsilon_\perp)} {k_BT c^2}\right)\exp\left(\frac{\gamma_Bv_{B,x'} p_{x'}}{k_BT}\right),\label{eq:dist2}
\end{eqnarray}
Plugging the equation into (??), we find that the this gives an associated growth rate of
\begin{equation}
\Gamma \simeq -\omega_p\frac{\pi n_b}{2 n_t}   m_e v_B^2 R,
\end{equation}
which involves the integral of 
\begin{equation}
R\equiv \int \frac{k\partial F_b/\partial p_{z'}}{\omega - k_{z'}v_{bz'}}d^3p,
\end{equation}
where $R$ is the residue for $p_{z'}$ such that $v_{z'} = v_{\rm ph}$.  
   Here we find for the two elements that contribute to the pole:
\begin{equation}
\left.
\frac{\partial}{\partial p_{z'}} \left(\omega-k v_{bz'}\right)
\right|_{\rm pole}
=
\left.
- k \left(
  \frac{c^2}{\epsilon} - \frac{p_{z'}^2 c^4}{\epsilon^3}
\right)
\right|_{\rm pole}
=
- \frac{k c^2}{\gph^3 \epsilon_\perp}\,,
\end{equation}
and 
\begin{equation}
\begin{aligned}
\left.
k\frac{\partial F_b}{\partial p_{z'}}
\right|_{\rm pole}
%&=
&=
\left.
-\frac{k\gamma_B}{k_BT} \left( \frac{p_{z'} c^2}{\epsilon} - v_{B,z'}\right)
F_b
\right|_{\rm pole}\\
&=
-
\frac{
  k
  (\vph-v_{B,z'}) \mu m_e^{-3}c^{-3}}{
  4\pi k_BT K_2(\mu)
} \exp\left(-\frac {\gamma_B\gph(c^2 - v_{B,z'}v_{\rm ph})\epsilon_\perp)} {k_BT c^2}\right)\\
&\exp\left(\frac{\gamma_Bv_{B,x'} p_{x'}}{k_BT}\right).
\end{aligned}
\end{equation}
%where we have used
%\begin{equation}
%\left.
%\gamma_b(\epsilon-v_{bz} p_z -v_{bx} p_x)/T
%\right|_{\rm pole}
%=
%\gamma_b \gph ( 1-v_{bz}\vph ) \epsilon_\perp/T - \gamma_b v_{bx} p_x/T\,.
%\label{eq:Eopole}
%\end{equation}

@@

Thus, we define $p_{x'}'=\gamma_w(p_{x'}-w \epsilon_\perp)$ and $p_{y'}'=p_{y'}$ and find:
\begin{eqnarray}
%\epsilon'_\perp &=& \gamma_w(\epsilon_\perp - w p_x),\\
\epsilon_\perp &=& \gamma_w(\epsilon'_\perp + w p'_x),\\
d p_x d p_y &=& \frac{\epsilon_\perp}{\epsilon'_\perp} d p'_x d p'_y.
\end{eqnarray}
Inserting this into equation (\ref{eq:I-integral}), we find
\begin{equation}
\begin{aligned}
\kI &= 
\int d^2\!p'_\perp \frac{\epsilon_\perp^2}{\epsilon'_\perp} \exp\left(-\frac{\cG'\epsilon'_\perp}{k_B T}\right)\\
%&=
%\gamma_w^2 \int d^2\!p'_\perp \left(
 % \epsilon'_\perp + 2 w 
%\qc{p'_x}
%+ w^2 \frac{p'^2_x}{\epsilon'_\perp}
%\right) e^{-\cG'\epsilon'_\perp/m}\\
&=
\pi \gamma_w^2 \int_0^\infty dp'^2_\perp \left(
  \epsilon'_\perp + \frac{w^2}{2} \frac{p'^2_\perp}{\epsilon'_\perp}
\right) \exp\left(-\frac{\cG'\epsilon'_\perp}{k_B T}\right)\,,
\end{aligned}
\end{equation}
where $\cG'\equiv\cG/\gamma_w$. Note in the second line that we have used isotropy in $\bp_{\perp}'$ to eliminate terms linear in $\bp_{\perp}'$. 
% and in the second line we used the fact
%that  $\epsilon'_\perp e^{-\cG'\epsilon'_\perp/m}$ is isotropic in
%$\bp'_\perp$.  
Using the following integrals: 
\begin{equation}
\int_0^\infty dx \sqrt{1+x} e^{-a\sqrt{1+x}} = \frac{2 e^{-a}}{a^3}\left(a^2 + 2a + 2\right)
\end{equation}
and 
\begin{equation}
\int_0^\infty dx \frac{x}{\sqrt{1+x}} e^{-a\sqrt{1+x}}
=
\frac{4(a+1)}{a^3} e^{-a}\,,
\end{equation}
we find
\begin{equation}
\kI =
\frac{2 \pi \gamma_w^2 m_e^3c^4}{\cG'^3\mu^3} \left[
\left(\cG'^2\mu^2+2\cG'\mu+2\right)
+
\frac{w^2}{2c^2}
\left(2\cG'\mu+2\right)
\right]\exp\left(-\cG'\mu\right).
\end{equation}
Inserting this into (\ref{eq:Rogen}) yields
\begin{equation}
R = \frac{
  \gph^3 \gamma_w^2 (\vph-v_{B,z'}) 
}{ 2 \mu m_ec^2 \cG'^3 K_2(\mu)}
\left[
\left(\cG'^2\mu^2 + 2\cG'\mu + 2\right)
+
\frac{\gamma_B^2 v_{B,x'}^2}{2\cG^2 c^2} \left(2 \cG'\mu + 2\right)
\right] \exp\left(-\cG'\mu\right)
\end{equation}
and therefore,
\begin{equation}
\Gamma \simeq - \Gamma_0
\frac{\pi \gamma_w^2 \gph^3 m (\vph-v_{B,z'})}{4 \gamma_b \mu K_2(\mu) \cG'^3}
\left[
\left( \cG'^2\mu^2 + 2\cG'\mu + 2 \right) 
+
\frac{\gamma_B^2 v_{B,x'}^2}{2\cG^2c^2} \left(2 \cG'\mu+ 2\right)
\right]
\exp(-\cG'\mu),
\label{eq:OGgen}
\end{equation}
where $\Gamma_0 \equiv \omega_p \gamma_B (n_b/n_t) (m_e v_B^2/k_B T)$ is the typical maximum growth rate.
%The above differs from the beam plasma instability only due to the
%minor difference in the definition of $\cG'$ and the second term.  When
%$v_{bx}=0$, it is precisely the beam plasma instability growth rate.

@@

%However, obtaining asymptotic expansions for the growth rate in the high and
%low temperature limits that can be easily solved for the maximum
%growth rate is difficult.  Even with the tremendous numerical insight
%provided by Figure \ref{fig:ObliqueGMZ}, it is unclear how to make
%significant progress.  However, some general remarks can be made.  It
%appears at the maximum growth rate, we have to leading order:
%\begin{equation}
%\begin{gathered}
%\vph\simeq\sqrt{2\delta Z_x}\,,\quad
%v_{bx}=(1-\delta Z_x) v_b\,,\quad
%v_{bz}\simeq\sqrt{2\delta Z_x}\,v_b\,,\quad\\
%v_{bz}-\vph\simeq\left\{
%\begin{aligned}
%&\frac{\sqrt{T/m}}{\gamma_b} & T\ll m\\
%&\frac{1}{2\gamma_b} & T\gg m
%\end{aligned}
%\right.\,,\quad
%\gph\simeq1+\delta Z_x\,,\quad
%\cG' \sim \frac{m}{T}\,,\\
%\text{and}\quad w \sim 1\,,
%\end{gathered}
%\end{equation}
%\qc{[can you add more details which equations you approximate under which conditions? e.g., I get $1/2\gamma_b^2$ for the hot beam case \ldots]}\\
%where $\delta Z_x\equiv 1 - v_{bx}/v_b$.  Inserting these into Equation
%(\ref{eq:OGgen}) suffices to show that there exists growth rates that
%are $\propto \Gamma_0$, with

%\begin{equation}
%\Gamma_M
%\simeq
%\left\{
%\begin{aligned}
%& 0.38 \Gamma_0 & T\ll m\\
%& 0.34 \Gamma_0 & T\gg m
%\end{aligned}
%\right.\,,
%\end{equation}
%for $\gamma_b\gtrsim10$, as seen in Figure (\ref{fig:ObliqueGMZ}).  
%
